---
title: "Homework #4"
format: html
author: Yefrid Cordoba
---

## Problem 1

| Source    | DF  | SS      | MS     | F     | P         |
|-----------|-----|---------|--------|-------|-----------|
| Treatment | 4   | 1010.56 | 252.64 | 29.84 | 3.545e-08 |
| Block     | 5   | 323.375 | 64.765 | 7.639 | 3.72e-04  |
| Error     | 20  | 169.33  | 8.467  |       |           |
| Total     | 29  | 1503.71 |        |       |           |

### a

```{r}
MSE <- 169.33/20
MSE
dft <- 4
dfE <- 20
dfb <- dfE/dft
dfb
MSb <- 64.675
SSb <- MSb * dfb
SSb
SSt <- 1010.56
MSt <- SSt / dft
MSt
Fb <- MSb / MSE
Fb
Ft <- 29.84
p_t <- pf(Ft, dft, dfE, lower.tail = F)
p_t
p_b <- pf(Fb, dfb, dfE, lower.tail = F)
p_b
```

### b 

```{r}
b <- dfb + 1
b
```
There were used `6` blocks in this experiment.

### c

$y_{ij} = \mu + \alpha_i + \beta_j + \epsilon_{ij}$

Where:

$i = 1, 2, ...,5$ number of treatments\
$j = 1, 2, ...,6$ number of blocks\
$y_{ij} =$ the observation in the $j^{th}$ block receiving the $i^{th}$ treatment\
$\mu =$ overall mean\
$\alpha_i =$ the effect of the $i^{th}$ treatment. $\sum_{i=1}^n \alpha_i = 0$\
$\beta_j =$ the effect of the $j^{th}$ block. $\sum_{i=1}^n \beta_i = 0$\
$\epsilon_{ij} =$ random error $\sim N(0, \sigma^2)$

### d

Based on the P-values for the treatment and the 