---
title: "Homework #4"
format: html
author: Yefrid Cordoba
---

## Problem 1

| Source | DF | SS | MS | F | P |
|----|----|----|----|----|----|
| Treatment | 4 | 1010.56 | [252.64]{.underline} | 29.84 | [3.545e-08]{.underline} |
| Block | [5]{.underline} | [323.375]{.underline} | 64.765 | [7.639]{.underline} | [3.72e-04]{.underline} |
| Error | 20 | 169.33 | [8.467]{.underline} |  |  |
| Total | 29 | 1503.71 |  |  |  |

### a

```{r}
MSE <- 169.33/20
MSE
dft <- 4
dfE <- 20
dfb <- dfE/dft
dfb
MSb <- 64.675
SSb <- MSb * dfb
SSb
SSt <- 1010.56
MSt <- SSt / dft
MSt
Fb <- MSb / MSE
Fb
Ft <- 29.84
p_t <- pf(Ft, dft, dfE, lower.tail = F)
p_t
p_b <- pf(Fb, dfb, dfE, lower.tail = F)
p_b
```

### b

```{r}
b <- dfb + 1
b
```

There were used `6` blocks in this experiment.

### c

$y_{ij} = \mu + \alpha_i + \beta_j + \epsilon_{ij}$

Where:

$i = 1, 2, ...,5$ number of treatments\
$j = 1, 2, ...,6$ number of blocks\
$y_{ij} =$ the observation in the $j^{th}$ block receiving the $i^{th}$ treatment\
$\mu =$ overall mean\
$\alpha_i =$ the effect of the $i^{th}$ treatment. $\sum_{i=1}^n \alpha_i = 0$\
$\beta_j =$ the effect of the $j^{th}$ block. $\sum_{i=1}^n \beta_i = 0$\
$\epsilon_{ij} =$ random error $\sim N(0, \sigma^2)$

### d

Based on the P-values for the treatment and assuming that the $\alpha = 0.05 >> 3.545e-08$, It can be concluded that the treatments are statistically significant, and have influence on the response variable of the block experiment performed.

## Problem 2

```{r}
batch <- factor(rep(1:5, each = 3))
treatment <- factor(rep(c("A", "B", "C"), times = 5)) 
strength <- c(52, 60, 56, 47, 55, 48, 44, 49, 45, 51, 52, 44, 42, 43, 38) 
data <- data.frame(batch, treatment, strength)
```

### a

For this randomized complete block design it is recognize that there are 3 drying processes ($a = 3$ treatments), and 5 batches ($b = 5$ blocks). A suitable model for this experiment is shown below:

$y_{ij} = \mu + \alpha_i + \beta_j + \epsilon_{ij}$

Where:

$i = 1, 2, 3$ number of treatments\
$j = 1, 2, 3, 4, 5$ number of blocks\
$y_{ij} =$ the observation in the $j^{th}$ block receiving the $i^{th}$ treatment\
$\mu =$ overall mean\
$\alpha_i =$ the effect of the $i^{th}$ treatment. $\sum_{i=1}^n \alpha_i = 0$\
$\beta_j =$ the effect of the $j^{th}$ block. $\sum_{i=1}^n \beta_i = 0$\
$\epsilon_{ij} =$ random error $\sim N(0, \sigma^2)$

### b

To evaluate the if there is a statistical difference between the treatments, it is necessary to produce an ANOVA table including the blocks and analyze the p-value of the treatment

```{r}
model <- aov(strength~treatment+batch, data = data)

summary(model)
```

Based on the p-value(0.014) that is smaller that our assumed $\alpha = 0.05$, we have strong evidence to conclude that at least one of the treatment's mean strength is different than the others.

### c

```{r}
TukeyHSD(model, "treatment")
```

Based on the Tukey's HSD comparisson, treatment B differs from treatment A and B. Treatment C does not differ from treatment A, considering that the range that is evaluated is not including 0, and the $p_{adj}-value$ is lower than our $\alpha = 0.05$ for the intervals that are statistically different.

### d

```{r}
plot(model, which = 1)
```

The residuals vs fitted plot, shows that based on the random position of all the data across the the graph, we can conclude that the data meet the equal variance assumption(not funnel shape). Also as the data does not follow a curved pattern, we can conclude that there is not presence of a non-linear relationship in the data.

### e

```{r}
plot(model, which = 2)
```

The normal Q-Q plot shows that the data follows a linear trend that is close to the theoretical normal distributed residuals, therefore, we can assume that the samples are from a normally distributed population.

## Problem 3 

```{r}
day <- factor(rep(c("Monday", "Tuesday", "Wednesday", "Thursday"), each = 3)) 
operator <- factor(c("A", "B", "C", "A", "C", "D", "B", "C", "D", "A", "B", "D")) 
kelvin <- c(780, 820, 800, 950, 920, 940, 880, 880, 820, 840, 780, 820) 
data1<-data.frame(day, operator, kelvin)
data1
```

### a

For this randomized incomplete block design it is recognize that there are 4 days ($b = 4$ blocks), and 3 operators for each day from a total of 4 operators ($a = 4$ treatments). A suitable model for this experiment is shown below:

$y_{ij} = \mu + \alpha_i + \beta_j + \epsilon_{ij}$

Where:

$i = 1, 2, 3, 4$ number of treatments\
$j = 1, 2, 3, 4$ number of blocks\
$y_{ij} =$ the observation in the $j^{th}$ block receiving the $i^{th}$ treatment\
$\mu =$ overall mean\
$\alpha_i =$ the effect of the $i^{th}$ treatment.\
$\beta_j =$ the effect of the $j^{th}$ block.\
$\epsilon_{ij} =$ random error $\sim N(0, \sigma^2)$

### b

```{r}
model1 <- lm(kelvin~day+operator, data = data1)
anova(model1)
```

Based on the P-value for the treatment effect, as it is larger than $\alpha = 0.05$, we have weak evidence to conclude that the operator has an effect in the screen color difference. 

## Problem 4

```{r}
tire <- factor(rep(1:4, each = 3)) 
compound <- factor(c("A", "B", "C", "A", "B", "D", "A", "C", "D", "B", "C", "D")) 
wear <- c(238, 238, 279, 196, 213, 308, 254, 334, 367, 312, 421, 412)
data2 <-data.frame(tire, compound, wear)
data2
```

### a

```{r}
model2 <- aov(wear~tire+compound, data = data2)
summary(model2)
```

$H_o : \mu_a = \mu_b = \mu_c = \mu_d$\
$H_a :$ at least one $\mu_i \neq \mu_j$ for $i \neq j$\

We have to test if $F_{obs} \geq F_{crit}$ to reject the null hypothesis.\
$P(F_{df_{tr}, df_{E}} \geq F_{obs} )$\
Our $\alpha = 0.05$ and having that this is larger than our p-value (0.00335), we reject the null hypothesis, and have strong evidence to conclude that at least one of the means is different, therefore, the compound used to make the tire have an effect on the wear of the tire.

### b

```{r}
#| message: false
#| warning: false
library(tidyverse)

data2 |>
  filter(compound == "A") |>
  pull(wear)|>
  t.test()

```

The 95% confidence interval for the mean wear for compound A is (154.92, 303.75)

### c

```{r}
#| warning: false
TukeyHSD(model2)
plot(TukeyHSD(model2))
plot(model2$fitted, model2$residuals)
```

Base on the 95% family-wise confidence interval using Tukey's HSD coefficient, the only two comparisons that are not significantly different are between compound A and B, and compounds C and D. All the other pair-wise comparisons are significantly different.

```{r}
y_hat <- data2 |>
  group_by(compound) |>
  summarise(mean = mean(wear))
y_hat <- data.frame(y_hat)
y_hat
data2|>
  group_by(tire) |>
  summarise(sum = sum(wear))
Ba <- 755 + 717 + 955
Bb <- 755 + 717 + 1145
Bc <- 755 + 955 + 1145
Bd <- 717 + 955 + 1145

B_comp <- factor(c("A", "B", "C", "D")) 
Total <- c(Ba, Bb, Bc, Bd)
Bt <- data.frame(B_comp, Total)
Bt
k <- 3
y_dot <- data2 |>
  group_by(compound) |>
  summarise(sum = sum(wear))
y_dot <- data.frame(y_dot)
y_dot

kyB = k*y_dot$sum - Bt$Total
kyB <- c(kyB)
kydf <- data.frame(B_comp, kyB)
kydf
t <- 4
r <- 3
lmd <- r*(k-1)/(t-1)
lmd
y_grand <- mean(y_hat$mean)
mu_hat <- y_grand + (kydf$kyB / (t * lmd))
Est_tmean <- data.frame(B_comp, y_hat$mean, kyB, mu_hat)
Est_tmean
MSE<- 350
W <- (qtukey(0.05, t, 5) / sqrt(2)) * sqrt ((2 * k * MSE)/(t * lmd))
W
```

